Fernando LOOCV

### 2016-10-21 reports/na_fraction_pheno_predability.Rmd
remove code boxes from report

# Prediction with Imputed mRNAs
```{r message = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr")
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/",
                     inline = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE, message = FALSE)
```

```{r DMY, cache = TRUE, dependson = "Rename-Traits"}
dmy <- loocv_clean %>%
  mutate(Dent_NA_Fraction = as.numeric(as.character(Dent_NA_Fraction)),
         Flint_NA_Fraction = as.numeric(as.character(Flint_NA_Fraction))) %>%
  filter(Trait == "DMY", Predictor == "mRNA") %>%
  droplevels() %>%
  unite(NA_Fraction, Dent_NA_Fraction, Flint_NA_Fraction, sep = "_") %>%
  group_by(Predictor, Trait, NA_Fraction) %>%
  summarize(Pred_Ability = cor(y, yhat)) %>%
  ungroup() %>%
  arrange(desc(Pred_Ability)) %>%
  separate(NA_Fraction,
           into = c("Dent_NA_Fraction", "Flint_NA_Fraction"), 
           sep = "_") %>%
  mutate(Dent_NA_Fraction = as.numeric(Dent_NA_Fraction),
         Flint_NA_Fraction = as.numeric(Flint_NA_Fraction)) %>%
  rowwise() %>%
  mutate(Avg_NA = mean(c(Dent_NA_Fraction, Flint_NA_Fraction)))
```

>>>>

# Prediction with Imputed mRNAs
```{r message = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr")
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE, message = FALSE)
```

```{r DMY, cache = TRUE, dependson = "Rename-Traits"}
dmy <- loocv_clean %>%
  mutate(Dent_NA_Fraction = as.numeric(as.character(Dent_NA_Fraction)),
         Flint_NA_Fraction = as.numeric(as.character(Flint_NA_Fraction))) %>%
  filter(Trait == "DMY", Predictor == "Transcriptomic") %>%
  droplevels() %>%
  unite(NA_Fraction, Dent_NA_Fraction, Flint_NA_Fraction, sep = "_") %>%
  group_by(Predictor, Trait, NA_Fraction) %>%
  summarize(Pred_Ability = cor(y, yhat)) %>%
  ungroup() %>%
  arrange(desc(Pred_Ability)) %>%
  separate(NA_Fraction,
           into = c("Dent_NA_Fraction", "Flint_NA_Fraction"), 
           sep = "_") %>%
  mutate(Dent_NA_Fraction = as.numeric(Dent_NA_Fraction),
         Flint_NA_Fraction = as.numeric(Flint_NA_Fraction)) %>%
  rowwise() %>%
  mutate(Avg_NA = mean(c(Dent_NA_Fraction, Flint_NA_Fraction)))
```

### 2016-10-21 reports/transcriptomic_data.Rmd
remove code boxes from report

# Transcriptomic Data
Keep only genotypes for which both, transcriptomic and genomic data are 
available.
```{r echo = FALSE}
pacman::p_load("sspredr", "tidyverse", "caret", "e1071", "viridis")
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", small.mark = ".", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/", inline = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
                      warning = FALSE, message = FALSE)
```

```{r mRNA-Data, cache = TRUE}
# Common genotypes
genos <- readRDS("./data/processed/common_genotypes.RDS")
dent <- as.character(genos$Dent$mrna)
flint <- as.character(genos$Flint$mrna)

# mRNA
mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna <- t(mrna)
mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
mrna_df <- mrna %>%
  as_data_frame() %>%
  mutate(G = rownames(mrna)) %>%
  mutate(Group = ifelse(G %in% dent, yes = "Dent", no = "Flint"),
         Group = as.factor(Group)) %>%
  select(-G)
```

@@

## Tidy-up data
```{r Tidy-Data, cache = TRUE}
mrna_lst <- list(Raw = mrna_lst, Trans = trans_lst)
tidy_mrna <- mrna_lst %>% 
  at_depth(.depth = 2, .f = ~as.data.frame(.)) %>%
  at_depth(.depth = 1, .f = ~bind_rows(., .id = "Group")) %>%
  bind_rows(.id = "Type") %>%
  as_tibble() %>%
  gather(key = "mRNA", value = "Expression", -Group, -Type) %>%
  mutate(Type = as.factor(Type),
         Group = as.factor(Group))
```

## Skewness
Compute the skewness and the Shapiro-Wilk test of normality for each mRNA
```{r Check-Normality, cache = TRUE}
mrna_skew <- tidy_mrna %>%
  group_by(Group, Type, mRNA) %>%
  summarize(Skewness = e1071::skewness(Expression),
            Shapiro = stats::shapiro.test(Expression)$statistic)
```

Plot the distribution of skewness values across all mRNAs
```{r, mRNA-Skewness, cache = TRUE}
mrna_skew %>%
  ggplot(aes(x = Skewness, fill = Type)) +
  geom_density(alpha = 0.2) +
  facet_grid(. ~ Group) +
  theme(legend.position = "top")

mrna_skew %>%
  ggplot(aes(x = Skewness, fill = Type)) +
  xlim(c(-2, 2)) +
  geom_density(alpha = 0.2) +
  facet_grid(. ~ Group) +
  theme(legend.position = "top")
```

Find the least skewed/non-normally distributed mRNAs.
```{r Association-Shapiro-Wilks-Skewness, cache = TRUE}
mrna_skew %>%
  mutate(Abs_Skewness = abs(Skewness)) %>%
  ggplot(aes(x = Shapiro, y = Abs_Skewness)) +
  xlim(c(0.8, 1)) +
  ylim(c(0, 2)) +
  geom_point() +
  facet_grid(Group ~ Type) +
  xlab("Shapiro-Wilks Test Statistic") +
  ylab("Absolute Skewness")
```

>>>>

# Transcriptomic Data
Keep only genotypes for which both, transcriptomic and genomic data are 
available.
```{r echo = FALSE}
pacman::p_load("sspredr", "tidyverse", "caret", "e1071", "viridis")
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", small.mark = ".", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
                      warning = FALSE, message = FALSE, echo = FALSE)
```

```{r mRNA-Data, cache = TRUE}
# Common genotypes
genos <- readRDS("./data/processed/common_genotypes.RDS")
mrna_inbreds <- genos %>%
  filter(Pool != "Hybrid", Data_Type == "mrna") %>%
  split(.$Pool) %>%
  map("G")

# mRNA
mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna <- t(mrna)
mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
mrna_df <- mrna %>%
  as_data_frame() %>%
  mutate(G = rownames(mrna)) %>%
  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
         Group = as.factor(Group)) %>%
  select(-G)
```

@@

```{r Tidy-Data, cache = TRUE}
mrna_lst <- list(Raw = mrna_lst, Trans = trans_lst)
tidy_mrna <- mrna_lst %>% 
  at_depth(.depth = 2, .f = ~as.data.frame(.)) %>%
  at_depth(.depth = 1, .f = ~bind_rows(., .id = "Group")) %>%
  bind_rows(.id = "Type") %>%
  as_tibble() %>%
  gather(key = "mRNA", value = "Expression", -Group, -Type) %>%
  mutate(Type = as.factor(Type),
         Group = as.factor(Group))
```

## Skewness
```{r Check-Normality, cache = TRUE}
mrna_skew <- tidy_mrna %>%
  group_by(Group, Type, mRNA) %>%
  summarize(Skewness = e1071::skewness(Expression),
            Shapiro = stats::shapiro.test(Expression)$statistic)
```

```{r, mRNA-Skewness, cache = TRUE, fig.cap = "Distribution of skewness values across all mRNAs, separately for transformed and raw BLUEs."}
mrna_skew %>%
  ggplot(aes(x = Skewness, fill = Type)) +
  xlim(c(-2, 2)) +
  geom_density(alpha = 0.2) +
  facet_grid(. ~ Group) +
  theme(legend.position = "top") +
  ylab("Density")
```

```{r Association-Shapiro-Wilks-Skewness, cache = TRUE, fig.cap = "Shapiro-Wilks Test Statistic vs absolute skewness for transcriptomic BLUEs depending on whether the BLUEs were transformed or not."}
mrna_skew %>%
  mutate(Abs_Skewness = abs(Skewness)) %>%
  ggplot(aes(x = Shapiro, y = Abs_Skewness)) +
  xlim(c(0.8, 1)) +
  ylim(c(0, 2)) +
  geom_point() +
  facet_grid(Group ~ Type) +
  xlab("Shapiro-Wilks Test Statistic") +
  ylab("Absolute Skewness")
```
